Tidy data structure to support exploration and modeling of temporal-context data

### 2018-11-12 main.Rmd
restarted time-based pipeline

@@

@@

Since Table \@ref(tab:tb-sub) possesses complete observations for each category, the interval obtained from the greatest common factor is 1 year.

Extensible. Custom index classes can be supported.

@@

What is notable about data pipeline is that it not only frees users from a tedious and error-prone analysis but also empowers a wider audience to focus on the statistical analysis tasks at hand without concerning the details of computational implementation. A fluent and coherent pipeline glues the grammar of data manipulation and tidy data as a fundamental unit of data analysis together. It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes as many as possible.

@@

This "tidy" data representation choice will prove suitable for almost all conceivable applications of temporal data. But the data abstraction cannot live alone with an additional toolchain in order to be able to acquire insight and knowledge about the data itself. A desirable toolbox chains transformation, visualisation, and modelling as a data science pipeline. Each component itself is also a self-contained pipeline that can be decomposed into many individual operations. This paper will focus on part of the general pipeline---data transformation, for example filtering observations, selecting, and summarising variables, by contrast to data cleaning. We shall see how this data abstraction breeds a consistent data process.

All the individual operations strive to keep the time context so that they will implicitly preserve the index variable.

@@

* **row-wise**: `filter()`, `slice()`, `arrange()`, `fill_na()`
* **column-wise**: `mutate()`, `select()`, `summarise()`
* **group-wise**: `index_by()`, `group_by()`
* **reshape**: `gather()`, `spread()`
* **time-wise**: `lead()`, `lag()`, `difference()`
* **rolling window**: `slide()`, `tile()`, `stretch()`

>>>>

```{r pipeline-ann}
include_graphics("img/time-series-structure.pdf")
```

@@

Additional representations are provided, like year-month, year-quarter. Extensible. Custom index classes can be supported.

@@

@@

The tidy data abstraction lays the foundation for data science modules including transformation, visualisation and modelling. Each module frictionlessly communicates between each other. Because a desirable toolbox acquires tidy input, produces tidy output, and consequently chains a series of operations together to accomplish analytic tasks at hand.

What is notable about data pipeline is that it not only frees users from a tedious and error-prone analysis but also empowers a wider audience to focus on the statistical analysis tasks at hand without concerning the details of computational implementation. A fluent and fluid pipeline glues tidy data and the grammar of data manipulation together. It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes as many as possible.

@@

This tidy data representation choice will prove suitable for almost all conceivable applications of temporal data. Time-based pipeline starts with raw data. It is advised to scrutinize identical entries of key and index in the right beginning. Duplicates signal the data quality issue, which would likely affect the following analyses and hence decision making. Analysts are encouraged to gaze at data as early as possible and reason about the process of data cleaning. When the data meets the tsibble standard, it flows into the exploration stage.

Time series models typically assume that the input is a complete and regularly-spaced series, which most temporal data can barely satisfy. Temporal data becomes more granular in time resolution and more disaggregated to individual level. This inevitably goes with implicit missing values and noisiness. Data needs transformation to some extent for modelling. A suite of verbs are introduced to flatten the lumpy path from temporal data to the object that directs to modelling under the umbrella of tsibble structure.

All the individual operations strive to keep the time context so that they will implicitly preserve the index variable, and maintain a valid tsibble all the time.

@@

* **implicit missingness**: `has_gaps()`, `count_gaps()`, `fill_na()`
* **row-wise**: `filter()`, `slice()`, `arrange()`
* **column-wise**: `mutate()`, `select()`, `summarise()`
* **group-wise**: `index_by()`, `group_by()`
* **reshape**: `gather()`, `spread()`, `nest()`, `unnest()`
* **rolling window**: `slide()`, `tile()`, `stretch()`
